---
title: "Computer Lab 1"
author: "Thomas Zhang"
date: "Tuesday, October 27, 2015"
output:
  pdf_document:
    fig_height: 4
    fig_width: 7
---


## Assignment 1

### 1.1 - 1.2
The spambase data set contains observations of e-mails with columns consisting of the frequencies of 48 words as well as an indicator of whether the observation is a Spam e-mail or not. To start, the spambase data set is divided into a train and a test data set. Each data set contains half of the original data set, where the observations have been randomly divided between the train data set and the test data set.
```{r, echo=FALSE,message=FALSE}
library(XLConnect)
library(kknn)

#You gonna need java 64-bit for rJava and XLConnect to work
#FROM "THIS FILE LOCATION", EXCEL FILES SHOULD BE FOUND IN A SUBFOLDER IN "THIS FILE LOCATION" CALLED DATA
wb = loadWorkbook("D:/R_HW/ML-lab-1/data/spambase.xlsx")
data = readWorksheet(wb, sheet = "spambase_data", header = TRUE)

n=dim(data)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.5))
train=data[id,]
test=data[-id,]
```

We wish to pretend as if we did not know whether the observations in the test data set are spam or not spam and that we have to classify them with a k-nearest-neighbor classifier. In order to do this, we use a distance function which, after we first scale the observations, takes into account the similarity of frequencies of words between two e-mail observations by way of calculating the dot product of two rows of the spambase data set (excluding the last column). Thus, the dot/cosine product distances between the observations are calculated and used to implement a k-nearest-neighbor classifier, called *knearest*. The distance function is given by 1 - c(X,Y) where c(X,Y), the cosine product, is defined as:  
$$ X^T Y \over \sqrt{\sum_{i=1}X_i^2}\sqrt{\sum_{i=1}Y_i^2} $$
In the spambase data set, there are rows with only zeros in all frequencies of words. these observations lie equidistant from all others and we simply delete them (while keeping in mind which rows they were).
The code for the knearest function can be seen in the appendix. 

```{r, echo=FALSE}
knearest <- function(data, k, newdata){
  
  dataspamindicator <- data$Spam
  data <- as.matrix(data)
  data <- data[,-ncol(data),drop=FALSE]


  # find out which data rows are the zero vector
  datazerorows <-rep(TRUE,nrow(data))
  for(g in 1:nrow(data)){
    if(sum(data[g,]) == 0){
      datazerorows[g] <- FALSE
    }  
  }
  
  data <- data[datazerorows,]
  dataspamindicator <- dataspamindicator[datazerorows]
  
  newdata <- as.matrix(newdata)
  newdata <- newdata[,-ncol(newdata),drop=FALSE]
  
  # find out which newdata rows are the zero vector
  newdatazerorows <-rep(TRUE,nrow(newdata))
  for(p in 1:nrow(newdata)){
    if(sum(newdata[p,]) == 0){
      newdatazerorows[p] <- FALSE
    }  
  }

  newdata <- newdata[newdatazerorows,]

  nearness <- matrix(0,k, as.numeric(nrow(newdata)))
  
  for(j in 1:nrow(newdata)){
    distancesfromj <- rep(0,nrow(data))
    for(i in 1:nrow(data)){
      normalizing <- sqrt(sum(data[i,]^2)) * sqrt(sum(newdata[j,]^2)) 
      result <- newdata[j,] %*% data[i,] / normalizing
      dist <- 1 - result
      
      distancesfromj[i] <- dist
    }
    distancesfromj <- sort(distancesfromj, index.return = TRUE)
    nearness[,j] <- distancesfromj$ix[1:k]
  }
  
  nearness <- t(nearness)
  predictedclassprobsfornewdata <- rep(0,nrow(nearness))
  for(m in 1:nrow(nearness)){
    predictedclassprobsfornewdata[m] <- sum(dataspamindicator[nearness[m,1:k]]) / k
  }
  
  return(list(predicted = predictedclassprobsfornewdata, deletedrows = newdatazerorows))
}

knn5 <- knearest(train,5,test)
knn1 <- knearest(train,1,test)
```

### 1.3-1.4
The implemented k-nearest-neighbor classifier returned the predicted probabilities of the test data observations being Spam e-mails, as well as which rows were deleted because they contained only zeros. The classification principle that decides how to classify each e-mail based upon its predicted probability of being Spam is stated as following:  

$\hat{y} = 1$ if p (Y = 1|X) > 0.5, otherwise $\hat{y} = 0$.

With K, the number of neighbors, set to 5 the confusion matrix generated from classifying the test data and the misclassification rate for the test data looks like this:  
```{r, echo=FALSE}
tabularize<-function(knn,decider,newdata){
  classified_by_knn <- rep(0,length(knn$predicted))
  for(u in 1:length(knn$predicted)){
    if(knn$predicted[u] > decider){
      classified_by_knn[u] <- 1
    }
  }
  testdataspam <-newdata$Spam[knn$deletedrows]
  
  tab <- table(Truth = testdataspam,classified_by_knn)
  return(tab)  
}

tab <- tabularize(knn5,.5,test)
print(tab)
paste("Misclassification rate:",round(1-sum(diag(tab))/sum(tab),4),sep=" ")
```

For the classification of the e-mails in the test data set using number of neighbours $K = 1$ we obtained confusion matrix and and the misclassification rate
```{r, echo=FALSE}
tab <- tabularize(knn1,.5,test)
print(tab)
paste("Misclassification rate:",round(1-sum(diag(tab))/sum(tab),4),sep=" ")
```

### 1.5
When the function *kknn* from the package **kknn** is used to classify the e-mails with the k-nearest-neighbor method, the following results are obtained. 
K is set to 5 and for the test data the confusion matrix looks like this:
```{r, echo=FALSE}
kknndata <- kknn( Spam ~. , train, test ,k = 5)

#str(kknndata)

tabularizekknn<-function(kknndata,decider){
  classified_by_kknn <- rep(0,length(kknndata$fitted.values))
  for(v in 1:length(kknndata$fitted.values)){
    if(kknndata$fitted.values[v] > decider){
      classified_by_kknn[v] <- 1
      }
  }
  truth <- test$Spam
  tab <- table(Truth = truth,classified_by_kknn)
  return(tab)
}


tab <- tabularizekknn(kknndata,0.5)
print(tab)
paste("Misclassification rate:",round(1-sum(diag(tab))/sum(tab),4),sep=" ")
```
When comparing with the *knearest* function it seems *knearest* performs slightly better than the *kknn* function. 

### 1.6
The sensitivity and specificity is computed for both the *knearest* and the *kknn* function. For each classification for the respective methods the sensitivity and specificity is computed according to the following formulas.  
$$ Sensitivity = \dfrac{True Positive}{All Positive} $$  
$$ Specificity = \dfrac{True Negative}{All Negative} $$

The ROC curve is the TPR plotted against the FPR where FPR = 1-Specificity.
The ROC for both models with $K=5$ are fitted and plotted in a graph.

```{r, echo=FALSE, fig.pos="center"}
#ROC is TPR tp/(tp +fn) against FPR fp/(fp + tn) FPR = 1 - Specificity
#Sensitivity is TPR, specificity is TNR tn/(tn + fp)

pi <- seq(0.05, 0.95, by=0.05)

collect<-function(pi,knn,kknndata,newdata){
    collector <-matrix(0,19,4)
    for(b in 1:length(pi)){
      target <- tabularize(knn,pi[b],newdata)
      target2 <- tabularizekknn(kknndata,pi[b])
      collector[b,1:2] <- c(target[2,2] / ( target[2,2] + target[2,1]) , target[1,1] / (target[1,1] + target[1,2]))
      collector[b,3:4] <- c(target2[2,2] / ( target2[2,2] + target2[2,1]) , target2[1,1] / (target2[1,1] + target2[1,2]))
      
    }
    return(collector)
}

plotthis <- collect(pi,knn5,kknndata,test)


kknn_ROC <- data.frame(cbind(1 - plotthis[,4], plotthis[,3]))
knearest_ROC <- data.frame(cbind(1 - plotthis[,2], plotthis[,1]))

plot(kknn_ROC, type="l", col="blue", xlim=c(0.02,0.3), ylim=c(0.6,0.97),
     xlab="1-Specificity (false positive rate)", ylab="Sensitivity (True positive rate)",
     main="ROC curve for kknn and knearest function")
lines(knearest_ROC, type="l", col="red")
legend(0.15,0.9,c("kknn","knearest"),
       lty=c(1,1),
       lwd=c(2.5,2.5),col=c("blue","red"))

```

Some conclusions about the ROC curves is that both models seem to be pretty good classifiers of spam and non-spam e-mails. There are no major differences between the models, but the knearest classifier seems to be a little more aggressive than the kknn classifier
as interpreted by the higher sensitvity it is capable of in return for having a lower possible specificity. A trade-off in these two characteristics seems likely. the ROC curve shows that when the decision value for the classification principle is low (high sensitvity) the *knearest* function performs better and for high decision values (low sensitivity) the *kknn* function performs a little bit better.

Looking more closely at the respective lines it can be seen that the red one, *knearest*, only change direction when the decision value reaches 0.2, 0.4, 0.6 and so on. That is because those are the only possible preditced probabilites when $K=5$ in *knearest*. Apparently there is some difference in how the classifiers use the information about the five nearest neighbors. 